Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-15 doc/manuscript/PNAS-template-supp-info-empty.tex

\dataset{gemma\_outliers.csv}{Description needed}

\dataset{spline\_window\_shared\_outliers.csv}{Description needed}

>>>>

\dataset{gemma\_outliers\_all.csv}{Set of liberal SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.05), identified via GEMMA.}

\dataset{gemma\_outliers\_strict\_freq.csv}{Set of strict SNP and INDEL outlier variants (Bonferroni corrected p-value $<$ 0.01), identified via GEMMA.
Additional information includes variant location within the gene (predicted\_gene\_loc) and whether the variant is synonymous or nonsynonymous (predicted\_effect\_AA).}

\dataset{spline\_window\_shared\_outliers.csv}{Description of overlapping \emph{F\textsubscript{ST}} and
\(\pi_{diff}\) outlier windows, as identified in the splined window
analysis.}

\dataset{spline\_window\_gene\_details.csv}{Annotation information for all genes within each of the
overlapping outlier windows in dataset S3}

### 2023-05-15 doc/manuscript/translocation_pnas_SI.tex

In the Results section of the main paper, we describe the stringent set
of outlier SNPS (identified using a Bonferroni-corrected p-value of
0.01). The liberal set, identified using a Bonferroni-corrected p-value
of 0.05, included 38 outliers (35 SNPs and 3 structural variants) from
30 distinct genes across 16 contigs.

Multiple genes in our stringent set of outliers have known associations
with amphibian Bd resistance/tolerance. Of particular interest is a
\textasciitilde940kb region on contig 8 that we identified as under
selection, and that contains 8 genes associated with either the MHC
Class I or Class III regions. The order of these genes in the MYL frog
genome is conserved from \emph{Xenopus} \citep{ohta2006} and is
consistent with MHC Class I/III linkage reported in the Komodo dragon
\citep{reed2021}. Numerous studies have linked MHC genes to frog
resistance/tolerance against Bd infection
\citep[e.g.,][]{savage2011, bataille2015}. In addition to the genes in
this region that we describe in the Results section of the main paper,
other genes also deserve mention. For example, we identified the MHC
Class III gene, heat shock protein 70 (HSP70), as occurring in this
region (Figure~\ref{fig-synteny-plot}). Similarly, \citep{byrne2020}
identified an HSP70-binding protein as a candidate gene under selection
in \emph{Atelopus varius} populations recovering after Bd-caused
declines. HSP70 was also highly expressed in Bd-infected versus
uninfected \emph{Xenopus tropicalis} \citep{rosenblum2009}. This MHC
gene cluster lies on the same chromosome as the outlier SNP in RIN3
(Fig. 2 C, Fig. 1
A).

Additionally, we identified a region on contig 19 that had the highest
\(\pi_{diff}\) of any outlier window, indicative of directional
selection. This region contained the gene, interferon-induced very large
GTPase 1-like (GVINP1). Interferons boost immune system processes and
previous studies of frogs have identified interferon-related genes as
associated with Bd resistance/tolerance \citep{byrne2020} or have shown
increased expression of this class of genes when frogs were exposed to
Bd in the lab \citep{rosenblum2009, ellison2014}. Variants in GVINP1
also explained approximately 20\% of the resistance to amoebic gill
disease in Atlantic salmon, a finding confirmed via expression analysis
\citep{robledo2018, robledo2020}. Furthermore, in \emph{Litoria
verreauxii alpina}, the related gene, IFN-induced very large GTPase,
showed lower expression in frogs from less susceptible populations
versus more susceptible populations \citep{grogan2018}.

Finally, we identified a region characterized by high
\emph{F\textsubscript{ST}} and low \(\pi_{diff}\) and that contained the
complement genes C6 and C7. This could indicate that balancing selection
is acting in this region of the genome to favor a diverse set of
alleles, as is known for C6 in humans \citep{soejima2005}.

@@

Finally, we assessed whether our stochastic model could reproduce
observed trajectories of population recovery. We focused on population
70550 because this was our longest CMR time series for a translocated
population and because this population shows evidence of substantial
post-translocation increases in adult abundance associated with
population establishment and recovery. We simulated our model for 16
years, repeating the simulation 50,000 times. For each run and each
year, we drew \(\omega\) from a uniform distribution between 0 and 1 (or
equivalently a beta distribution with mean 0.5 and \(\phi = 2\)). Using
Approximate Bayesian Computing and rejection sampling
\citep{kosmala2016}, we identified the top 2\% of trajectories (i.e.,
1000 trajectories) that minimized the sum of squared errors between the
observed and predicted data. The \(\omega\) values associated with these
``best'' trajectories represented an approximate posterior distribution
\citep{beaumont2010}. Using these best fit trajectories, we assessed
whether our model could qualitatively describe the patterns of recovery
in the observed data for population 70550.

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/pca_qq_manhattan.png}

@@

\caption{\label{fig-selectionresults}Results from the exome capture
selection study. (A) PCA calculated from binary SNPs showing genomic
relationship of samples. Numeric labels and colors match those in
Fig. 1 A. (B) qqplot showing observed and
expected p-values for 148,307 SNPs and indels as calculated in GEMMA.
Dashed line shows p-value that identifies outliers. (C) Manhattan plot
showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by
genomic position and contigs are sorted by size. Red line shows p-value
that identifies outliers. Outlier SNPs above this threshold are
highlighted and labeled. Bold labels indicate at least one
non-synonymous SNP is present in that gene.}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/synteny_figure.png}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/map_translocation_points.png}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/mcmc_areas_m2b.png}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/bdload_beforeafter.png}

@@

\caption{\label{fig-bdload-beforeafter}For frogs translocated to each of
the 12 recipient sites, Bd loads for the period immediately prior to
translocation versus during the 1-year period after translocation. Bd
loads are expressed as the number of ITS1 copies per skin swab, as
estimated by qPCR of the Bd ITS1 region. Loads indicative of severe
disease are \textgreater{} 5.8 ITS copies (on a log\textsubscript{10}
scale; see SI text for details). Box plot details are as in
Fig. 4.}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/mcmc_areas_m1d.png}

@@

\caption{\label{fig-survival-postdens}For the best model identifying
predictors of 1-year frog survival following translocation, estimated
posterior density curves and shaded 95\% uncertainty intervals for the
intercept and all predictor variables. In the Bayesian framework in
which the model was developed, variables are considered important
predictors if the associated uncertainty interval does not overlap zero
(indicated by the dashed red line).}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/compare_surv_probs.jpg}

@@

\caption{\label{fig-compare_surv_probs}Comparison of the average yearly
adult survival probabilities for adults translocated to each of three
sites versus adults that were naturally recruited at each site (as a
result of reproduction by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohort. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. The large error bars for 70550
reflect the large within and among individual variability in yearly
survival probability at this site (as also reflected in one-year
survival probabilities in Fig. 3). All
estimates are derived from the mrmr package.}

@@

{\centering \includegraphics[width=0.8\textwidth]{figures/pop_viability_figures_for_supp.jpg}

>>>>

In the Results section of the main paper, we describe the stringent set
of outlier SNPS (identified using a Bonferroni-corrected p-value of
0.01). The liberal set, identified using a Bonferroni-corrected p-value
of 0.05, included 38 outliers (35 SNPs and 3 INDELS) from 30 distinct
genes across 16 contigs. Our analysis found no significantly
over-represented GO terms in this set of 30 genes, or in the set of 38
outlier regions identified with the splined window analysis.

In the comparison of individual outlier variants between naive and
recovering populations, we detected consistent allele frequency changes
across all genetic clusters for 1 of 7 of the associated genes (RIN3;
Fig. 1). Nonetheless, for the other 6 genes
we still found a statistically significant signal of selection in 2 of
the 3 genetic clusters (containing populations 6, 7, 8, and 9;
Fig. 1, Figure~\ref{fig-selectionresults} A
SI). Therefore, although some outlier variant associations have a more
limited geographic extent than RIN3, they still describe parallel
evolutionary changes following Bd exposure and are therefore interesting
candidates for future studies.

@@

Finally, we assessed whether our stochastic model could reproduce
observed trajectories of population recovery. We focused on population
70550 because this was our longest CMR time series for a translocated
population and because this population shows evidence of substantial
post-translocation increases in adult abundance associated with
population establishment and recovery. We simulated our model for 16
years, repeating the simulation 50,000 times. For each run and each
year, we drew \(\omega\) from a uniform distribution between 0 and 1 (or
equivalently a beta distribution with mean 0.5 and \(\phi = 2\)). Using
Approximate Bayesian Computing and rejection sampling
\citep{kosmala2016}, we identified the top 2\% of trajectories (i.e.,
1000 trajectories) that minimized the sum of squared errors between the
observed and predicted data. The \(\omega\) values associated with these
``best'' trajectories represented an approximate posterior distribution
\citep{beaumont2010}. Using these best fit trajectories, we assessed
whether our model could qualitatively describe the patterns of recovery
in the observed data for population 70550. Code to replicate the
analyses can be found at \url{https://github.com/SNARL1/translocation}.

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/pca_qq_manhattan.png}

@@

\caption{\label{fig-selectionresults}Evidence for selection of
individual variants in recovering MYL frog populations. (A) PCA
calculated from binary SNPs showing the genomic relationship of samples.
Numeric labels and colors match those in
Fig. 1 A. (B) qqplot showing observed and
expected p-values for 148,307 SNPs and INDELS as calculated in GEMMA.
Dashed line shows p-value that identifies outliers. (C) Manhattan plot
showing p-values for each SNP as calculated by GEMMA. SNPs are sorted by
genomic position and contigs are sorted by size. Red line shows p-value
that identifies outliers. Outlier SNPs above this threshold are
highlighted and labeled. Bold labels indicate at least one
non-synonymous SNP is present in that gene.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/synteny_figure.png}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/map_translocation_points.png}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/mcmc_areas_m2b.png}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/bdload_beforeafter.png}

@@

\caption{\label{fig-bdload-beforeafter}Evidence that frogs translocated
from recovering populations can maintain the benefits of
resistance/tolerance in non-natal habitats. For frogs translocated to
each of the 12 recipient sites, Bd loads for the period immediately
prior to translocation versus during the 1-year period after
translocation. Bd loads are expressed as the number of ITS1 copies per
skin swab, as estimated by qPCR of the Bd ITS1 region. Loads indicative
of severe disease are \textgreater{} 5.8 ITS copies (on a
log\textsubscript{10} scale; see SI text for details). Box plot details
are as in Fig. 4.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/mcmc_areas_m1d.png}

@@

\caption{\label{fig-survival-postdens}Evidence that Bd load in not an
important predictor of post-translocation frog survival. For the best
model identifying predictors of 1-year frog survival following
translocation, estimated posterior density curves and shaded 95\%
uncertainty intervals for the intercept and all predictor variables. In
the Bayesian framework in which the model was developed, variables are
considered important predictors if the associated uncertainty interval
does not overlap zero (indicated by the dashed red line).}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/compare_surv_probs.jpg}

@@

\caption{\label{fig-compare_surv_probs}Evidence that frog
resistance/tolerance in non-natal habitats is maintained across
generations. Comparison of the average yearly adult survival
probabilities for adults translocated to each of three sites versus
adults that were naturally recruited at each site (as a result of
reproduction by translocated frogs). In contrast to
Fig. 3, these are not survival
probabilities from the first year following translocation, but instead
represent averaged survival probabilities across multiple years and
cohort. Points are median estimates and error bars give the 95\%
uncertainty intervals around the estimates, accounting for yearly
variation in survival probabilities. The large error bars for 70550
reflect the large within and among individual variability in yearly
survival probability at this site (as also reflected in one-year
survival probabilities in Fig. 3). All
estimates are derived from the mrmr package.}

@@

{\centering \includegraphics[width=0.85\textwidth]{figures/pop_viability_figures_for_supp.jpg}
